Slightly Positive Correlation Between Racialization, Gendering and Police Strip Searches in the Toronto Municipal, 2020-2021

### 2024-01-26 outputs/paper/paper.qmd
polish formatting and finalize paper

---
title: "Slightly Positive Correlation Between Racialization, Gendering and Police Strip Searches in the Toronto Municipal, 2020-2021"
title-block-banner: true
abstract: "Police strip search has been a topic of great contention during the COVID-19 pandemic. The question commonly raised was how certain demographics are more likely to undergo strip searches than others, and why has this skewed power dynamics been maintained for so long. Data then was gathered from Open Data Toronto to assess the correlation between strip searches and the racial and gender identities of arrested 'suspects' in the Toronto Municipal from 2020 to 2021. In short, this paper finds that police strip searches are slightly racialized and gendered, especially towards people of African descent and females within Toronto. This analysis would enable us to rethink the modern form of policing, how it transpires in times of crisis and critically affects people of marginalized identities." 
thanks: "Code and data are available at: https://github.com/ponolite/Police_Arrests_and_Strip_Searches/tree/main"
author: "Quang Mai"
date: today
date-format: long
format: pdf
editor: visual
toc: true
number-sections: true
bibliography: references.bib
output: 
  pdf_document:
    extra_dependencies: ["float"]
---

@@

This paper is organized into the following sections: Data, Results and Discussion. First, the Data section divulges the nature of the dataset obtained from Open Data Toronto and how data cleaning and extraction took place [@torontoPolice]. Then, all the trends and findings discovered are shown in the Results section. while the Discussion section further assesses these findings and summarizes the paper's main insights.

@@

Informative as it is, the dataset contains much naming inconsistency and many variables outside the scope of this paper. Thus, in order to conduct an intersectional---or gender and race---analysis of police strip searches, I have to further clean this dataset and split it into two smaller datasets. Additional details on data extraction and cleaning will be specified in subsequent sections.

@@

Most relevant to the paper are the variables on the racial and gender identities of people who were arrested and then strip-searched. To streamline this data, the names within the original dataset are simplified and then the dataset is split into a smaller dataset containing relevant columns ([see @tbl-cleaned_race_gender_data]). This dataset entails demographic and strip search-relevant information for each police arrest event, such as race, gender, whether the event involves a strip search, and if it is a strip search, are there any items found.

@@

To further inspect the nature of each strip search, reasons for strip searching prove a vital data source. Thus, this dataset--another sub-dataset split from the orginal raw data--conveys all police arrests that involve strip searching and the reasons behind strip searching. Under close examination, the naming of individual data cells from the original, raw dataset is quite disordered, for instance, using `None` and `XX` in place of the usual `NA` to differentiate between variables. Thus, all inconsistent data cells were renamed to streamline the naming conventions and data processing later on. After renaming, the big dataset was refined into containing relevant columns ([see @tbl-cleaned_search_reasons_data]). Then, for this particular dataset, all instances of police arrests without strip searching are dropped using the `filter()` function, along with a `!is.na()` function for all `reason` variables with `NA` as their values.

@@

To encapsulate, throughout 2020 and 2021, there are a total of 7801 recorded strip searches out of 65276 incidents of police arrests, which accounts for a whooping 12% of all arrests (see [@tbl-items-strip-search]).

@@

Overall, the data unfolds the relationship between strip searching and social identities such as race and gender. Within these intersecting identities, it was found that if there is an police arrest in Toronto from 2020 to 2021, strip searching would most likely happen to an Indigenous person or a black person. This is a statistically improbable insight, considering the major population make-up of white people in Toronto and their huge counts of arrests as seen in past bar graphs (see [@fig-race-gender-search]). Out of all the minority racial groups, at 2434 people, the black population has the highest counts of strip searches. Regarding reasons accounted for each strip search, minority racial groups are more likely to be strip searched without any reason provided or with less reasons provided compared to those that are ethnically white. Gender-wise, even though arrested males prevail in terms of being strip searched (see [@fig-race-gender-search]), females are more likely to be searched with less reasons provided. These insights all substantiate the assumptions that Toronto police strip search is racialized and gendered, which calls for a legal and political revisions of this intrusive form of policing.

>>>>

---
title: "Slightly Positive Correlation Between Racialization, Gendering and Police Strip Searches in the Toronto Municipal, 2020-2021"
title-block-banner: true
abstract: "Police strip search has always been a topic of great contention, where question commonly raised was how certain demographics are more likely to undergo strip searches than others. Data then was gathered from Open Data Toronto to assess the correlation between strip searches and the racial and gender identities of arrested 'suspects' in the Toronto Municipal from 2020 to 2021. In short, this paper finds that police strip searches are slightly racialized and gendered, especially towards people of African descent and females within Toronto. This analysis would enable us to rethink the modern form of policing, how it transpires in times of crisis and critically affects people of marginalized identities." 
thanks: "Code and data are available at: https://github.com/ponolite/Police_Arrests_and_Strip_Searches/tree/main"
author: "Quang Mai"
date: today
date-format: long
format: pdf
editor: visual
toc: true
number-sections: true
bibliography: references.bib
output: 
  pdf_document:
    extra_dependencies: ["float"]
---

@@

This paper is organized into the following sections: Data, Results and Discussion. First, the Data section divulges the nature of the dataset obtained from Open Data Toronto and how data extraction took place [@torontoPolice]. Then, all the trends and findings discovered are shown in the Results section. while the Discussion section further assesses these findings and summarizes the paper's main insights.

@@

Informative as it is, the dataset contains much naming inconsistency and many variables outside the scope of this paper. Thus, in order to conduct an intersectional---or gender and race---analysis of police strip searches, I have to further clean this dataset and split it into two smaller datasets.

@@

Most relevant to the paper are the variables on the racial and gender identities of people who were arrested and then strip-searched. There are 9 unique racial groups for the `race` variables and 3 unique gender groups for the `gender` variables. To streamline this data, the names within the original dataset are simplified and then the dataset is split into a smaller dataset containing relevant columns ([see @tbl-cleaned_race_gender_data]). This dataset entails demographic and strip search-relevant information for each police arrest event, such as race, gender, whether the event involves a strip search, and if it is a strip search, are there any items found.

@@

To further inspect the nature of each strip search, reasons for strip searching prove a vital data source. Thus, this dataset--another sub-dataset split from the orginal raw data--conveys all police arrests that involve strip searching and the reasons behind strip searching. Under close examination, the naming of individual data cells from the original, raw dataset is quite disordered, for instance, using `None` and `XX` in place of the usual `NA` to differentiate between variables. Thus, all inconsistent data cells were renamed to streamline the naming conventions and data processing later on. After renaming, the big dataset was refined into containing relevant columns ([see @tbl-cleaned_search_reasons_data]). Then, for this particular dataset, all instances of police arrests without strip searching are dropped using the `filter()` function.

@@

To encapsulate, throughout 2020 and 2021, there are a total of 7801 recorded strip searches out of 65276 incidents of police arrests, which accounts for 12% of all arrests (see [@tbl-items-strip-search]).

@@

Overall, the data unfolds the relationship between strip searching and social identities such as race and gender. Within these intersecting identities, it was found that if there is an police arrest in Toronto from 2020 to 2021, strip searching would most likely happen to an Indigenous person or a black person. This is a statistically improbable insight, considering the major population make-up of white people in Toronto and their huge counts of arrests as seen in past bar graphs (see [@fig-race-gender-search]). Out of all the minority racial groups, at 2434 people, the black population has the highest counts of strip searches. Regarding reasons accounted for each strip search, minority racial groups are more likely to be strip searched without any reason provided or with less reasons provided compared to those that are ethnically white. Gender-wise, even though arrested males prevail in terms of being strip searched (see [@fig-race-gender-search]), females are more likely to be searched with less reasons provided. These insights all substantiate the assumptions that Toronto police strip searches are racialized and gendered, which calls for a legal and political revisions of this intrusive form of policing as it hinges on racial biases, profiling, social intimidation and sexual invasion.
